---
title: "Advanced plots and inference"
subtitle: "SISBID 2017 <br> https://github.com/SISBID/Module2"
author: "Di Cook (dicook@monash.edu, @visnut) <br> Heike Hofmann (heike.hofmann@gmail.com, @heike_hh)"
date: "07/12-14/2017"
output:
  xaringan::moon_reader:
    css: ["default", "myremark.css"]
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      incremental: true
---

```{r echo = FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "",
  fig.height = 4,
  fig.width = 8,
  fig.align = "center",
  cache = FALSE
)
```

```{r echo=FALSE}
library(tidyverse)
library(ggthemes)
library(maps)
library(ISOcodes)
library(scales)
library(RColorBrewer)
library(gridExtra)
library(HLMdiag)
library(viridis)
```

# Tidy data and random variables

- The concept of tidy data matches elementary statistics
- Tabular form puts variables in columns and observations in rows
- Not all tabular data is in this form
- This is the point of tidy data

$$X = \left[ \begin{array}{rrrr}
           X_1 & X_2 & ... & X_p 
           \end{array} \right] \\
  = \left[ \begin{array}{rrrr}
           X_{11} & X_{12} & ... & X_{1p} \\
           X_{21} & X_{22} & ... & X_{2p} \\
           \vdots & \vdots & \ddots& \vdots \\
           X_{n1} & X_{n2} & ... & X_{np}
           \end{array} \right]$$

- $X_1 \sim N(0,1), ~~X_2 \sim exp(1) ...$
---
# Grammar of graphics and statistics

- A statistic is a function on the values of items in a sample, e.g. for $n$ iid random variates $\bar{X}_1=\sum_{i=1}^n X_{i1}$, $s_1^2=\frac{1}{n-1}\sum_{i=1}^n(X_{i1}-\bar{X}_1)^2$
- We study the behaviour of the statistic over all possible samples of size $n$. 
- The grammar of graphics is the mapping of (random) variables to graphical elements, making plots of data into statistics

---
# Pipeline: Messy to tidy to plot

```{r echo=TRUE}
messy_data <- read_csv("../data/tb.csv")
head(messy_data)
```

---

```{r echo=TRUE}
tidy_data <- messy_data %>% 
  gather(demo, count, -year, -iso2, na.rm = TRUE) %>% 
  separate(demo, c("gender", "age"))
tidy_data <- tidy_data %>% 
  filter(!(age %in% c("014", "04", "514", "u")))
head(tidy_data)
```

---
# 100% charts

```{r fig.width=10, fig.height=3, echo=TRUE}
tidy_data %>% 
  filter(iso2 == "US") %>% 
  ggplot(aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(~ age) +
  theme(
    axis.text = element_text(size = 14), 
    strip.text = element_text(size = 16), 
    axis.title = element_text(size = 16)
  )
```

---
# Stacked barcharts

```{r fig.width=10, fig.height=3, echo=TRUE}
tidy_data %>% 
  filter(iso2 == "US") %>% 
  ggplot(aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_grid(~ age) +
  theme(
    axis.text = element_text(size = 14), 
    strip.text = element_text(size = 16), 
    axis.title = element_text(size = 16)
  )
```

---
# Side-by-side barcharts

```{r fig.width=10, fig.height=3, echo=TRUE}
tidy_data %>% 
  filter(iso2 == "US") %>% 
  ggplot(aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity", position="dodge") +
  facet_grid(~ age) +
  theme(
    axis.text = element_text(size = 14), 
    strip.text = element_text(size = 16), 
    axis.title = element_text(size = 16)
  )
```

---
# facetted barcharts

```{r fig.width=10, fig.height=3, echo=TRUE}
tidy_data %>% 
  filter(iso2 == "US") %>% 
  ggplot(aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_grid(gender ~ age) +
  theme(
    axis.text = element_text(size = 14), 
    strip.text = element_text(size = 16), 
    axis.title = element_text(size = 16)
  )
```

---
# Rose plot

```{r fig.width=10, fig.height=3, echo=TRUE}
tidy_data %>% 
  filter(iso2 == "US") %>% 
  ggplot(aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_grid(gender ~ age) +
  theme(
    axis.text = element_blank(), 
    strip.text = element_text(size = 16), 
    axis.title = element_text(size = 16)
  ) + coord_polar()
```

---
# Rainbow charts

```{r fig.width=10, fig.height=3, echo=TRUE}
tidy_data %>% 
  filter(iso2 == "US") %>% 
  ggplot(aes(x = 1, y = count, fill = factor(year))) +
  geom_bar(stat = "identity", position="fill") +
  facet_grid(gender ~ age) +
  theme(
    axis.text = element_blank(), 
    strip.text = element_text(size = 16), 
    axis.title = element_text(size = 16)
  ) 
```

---
# Pie charts

```{r fig.width=10, fig.height=3, echo=TRUE}
tidy_data %>% 
  filter(iso2 == "US") %>% 
  ggplot(aes(x = 1, y = count, fill = factor(year))) +
  geom_bar(stat = "identity", position="fill") +
  facet_grid(gender ~ age) +
  theme(
    axis.text = element_blank(), 
    strip.text = element_text(size = 16), 
    axis.title = element_text(size = 16)
  ) + coord_polar(theta="y")
```

---
class: inverse middle 
# Your turn

What do you learn about tb incidence in the USA by gender and age and year from the
- 100% charts?
- stacked bar charts?
- side-by-side barcharts?
- facetted barcharts?

```{r eval=FALSE, echo=FALSE}
The US has been rocking it at reducing tb incidence, across most age groups, and gender. Different focus using the different charts:
  - 
  - 
  - 
  - 
```

---
# Maps

Make a world map, by merging iso2 with country codes to make a chloropleth

```{r}
data("ISO_3166_1")
world_map <- map_data("world")
tb_names <- tidy_data %>% 
  left_join(ISO_3166_1, by=c("iso2"="Alpha_2")) %>% 
  select(-Alpha_3, -Numeric, -Official_name, -Common_name) %>%
  rename(region=Name) %>%
  mutate(region=recode(region, 
                       "United States"="USA", 
                       "United Kingdom"="UK",
                       "Russian Federation"="Russia"))
tb_names_inc <- tb_names %>% 
  group_by(region) %>%
  nest() %>%
  mutate(
    model = purrr::map(data, ~ lm(count ~ year, data = .))
) %>%
  unnest(model %>% purrr::map(broom::tidy)) %>%
  select(region, term, estimate) %>% 
  spread(term, estimate)
```

---

```{r}
ggplot(tb_names_inc, aes(map_id = region)) + 
    geom_map(aes(fill=year), map = world_map, color="grey70", size=0.3) + 
    expand_limits(x = world_map$long, y = world_map$lat) +
    theme_few() +
    scale_fill_viridis("trend") + 
    theme(legend.position = "bottom",
         axis.ticks = element_blank(), 
         axis.title = element_blank(), 
         axis.text =  element_blank()) +
    guides(fill = guide_colorbar(barwidth = 15, barheight = .5))
```

---
# Inference

- Choice of plot implicitly sets $H_0$, $H_1$
- Generically, we are thinking  $H_0$: no pattern, $H_1$: pattern, but the choice of plot makes this much more explicit

---
# Putting the pieces together

```{r}
df <- data.frame(v1=c(rexp(20), rexp(15, 2)), 
                 v2=c(rep("A", 20), rep("B", 15)))
ggplot(df, aes(x=v2, y=v1, fill=v2)) + 
  geom_boxplot() + coord_flip()
```

---
class: inverse middle 
# Your turn

- Question?
- Data, variables
- Mapping
- Null generating mechanism

--
- Question? Is there a difference between the two groups? $H_0$: no difference, $H_1$: difference

--
- Data, variables: Two variables: v1, v2; v2 is categorical

--
- Mapping: x=V2, y=V1, colour=V1, geom=boxplot

--
- Null generating mechanism: permute the values of V1, relative to V2


---
# Clarity

- The null hypothesis is determined based on the plot type
- It is not based on the structure seen in a data set

---
# Lineup

Embed the data plot in a field of null plots

```{r eval=FALSE}
library(nullabor)
pos <- sample(1:20, 1)
df_null <- lineup(null_permute('v1'), df, pos=pos)
ggplot(df_null, aes(x=v2, y=v1, fill=v2)) + 
  geom_boxplot() +
  facet_wrap(~.sample, ncol=5) + coord_flip()
```

---

Which plot shows the most difference between the groups?

```{r fig.height=7, fig.width=9, echo=FALSE}
library(nullabor)
pos <- sample(1:20, 1)
df_null <- lineup(null_permute('v1'), df, pos=pos)
ggplot(df_null, aes(x=v2, y=v1, fill=v2)) + 
  geom_boxplot() +
  facet_wrap(~.sample, ncol=5) + coord_flip()
```

---
# Evaluation

- Computing $p$-values
- Power - signal strength

---
# p-values

Suppose $x$ individuals selected the data plot from a lineup of $m$ plots, shown to $K$ independent observers, then simplistically we can think about the probability of this happening, if the data plot is from the same distribution as the null plots. This yields a binomial formula:

$$P(X\geq x) = \sum_{i=x}^{K} \binom{K}{i} \left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{K-i}$$

For $x=4, K=17, m=20$

```{r}
pvisual(4, 17, m=20)
```

---
# Simulation approach

- Scenario I: in each of K evaluations a different data set and a different set of (m-1) null plots is shown.
- Scenario II: in each of K evaluations the same data set but a different set of (m-1) null plots is shown.
- Scenario III: the same lineup, i.e. same data and same set of null plots, is shown to K different observers.

---
# Simulation

Crucial idea: assign a p-value to each plot (data and null); under null hypothesis, this p-value is from U[0,1] 

Scenario I:
- for the $k$th lineup evaluation do:
 - pick 20 $p$-values from $U[0,1]$
 - for data plot compute 'strength' of other plots: $q = (1-p_\text{data})/\sum_j(1-p_j)$
 - Use $q$ to determine whether data was picked in simulation: $x_k \tilde B_{1,q}$
 - repeat above three steps $K$ times, and find the number of data picks $X = \sum_k x_k$
- Repeat N times to get distribution of $X$

---
# Simulation


Scenario II (same data, different nulls):
- for the $k$th lineup evaluation pick 20 $p$-values from $U[0,1]$:
- for data plot compute 'strength' of other plots: $q = (1-p_\text{data})/\sum_j(1-p_j)$
- Use $q$ to determine whether data was picked in simulation: $x_k \tilde B_{1,q}$
- find the number of data picks $X = \sum_k x_k$
- Repeat N times to get distribution of $X$

---
# Simulation


Scenario III (same data, same nulls):
- for the $k$th lineup evaluation pick $p_\text{data} \sim U[0,1]$:
 - pick 19 $p$-values from $U[0,1]$
 - for data plot compute 'strength' of other plots: $q = (1-p_\text{data})/\sum_j(1-p_j)$
 - simulate number of data picks $X ~ B_{K, q}$
- Repeat N times to get distribution of $X$


---
# Null-generating mechanisms

- Permutation: randomizing the order of one of the variables breaks association, but keeps marginal distributions the same
- Simulation: from a given distribution, or model. Assumption is that the data comes from that model 

---
class: inverse middle 
# Your turn

For these plot descriptions, decide on:

- null hypothesis
- null generating mechanism

---
class: inverse middle 
# Your turn

```{r echo=FALSE}
ggplot(autism, 
       aes(x=age2+2, y=vsae, group=childid, colour=gender)) + 
  geom_point() +
  geom_line() + xlim(c(0, 15)) +
  xlab("Age (in years)") + ylab("Vineland Socialization Age Equivalent")
```

---
class: inverse middle 
# Your turn

```{r echo=FALSE}
fly <- read_csv("../data/flying-etiquette.csv")
fly$`How often do you travel by plane?` <- 
  factor(fly$`How often do you travel by plane?`, levels=c(
    "Never","Once a year or less","Once a month or less",
    "A few times per month","A few times per week","Every day"))
fly_sub <- fly %>% 
  filter(`How often do you travel by plane?` %in% 
       c("Once a year or less","Once a month or less")) %>%
  filter(!is.na(`Do you ever recline your seat when you fly?`)) %>%
  filter(!is.na(Age)) %>% filter(!is.na(Gender))
ggplot(fly_sub, 
       aes(x=`In general, is itrude to bring a baby on a plane?`)) + 
  geom_bar(mapping=aes(fill=Gender), position="fill") + 
  coord_flip() 
```

---
class: inverse middle 
# Your turn

- For the tb data, subset to 2008
- Decide on two possible plot designs to show the relationship between age, gender.
- Write the code to make the plots
- Determine an appropriate null generating mechanism to test $H_o$: no relationship between age and gender.
- Make a lineup to for each plot design. 
- Show the lineups to 5 other participants, for each plot design
- Record the number of people selecting the data plot
- Compute the power for each test, the signal strength of the design



---
# Resources

- Hofmann, H., Follett, L., Majumder, M. and Cook, D. (2012) Graphical Tests for Power Comparison of Competing Designs, http://doi.ieeecomputersociety.org/10.1109/TVCG.2012.230.
- Wickham, H., Cook, D., Hofmann, H. and Buja, A. (2010) Graphical Inference for Infovis,  http://doi.ieeecomputersociety.org/10.1109/TVCG.2010.161. 

---
# Share and share alike

This work is licensed under the Creative Commons Attribution-Noncommercial 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc/3.0/us/ or send a letter to Creative Commons, 171 Second Street, Suite 300, San Francisco, California, 94105, USA.
