---
title: "Tidying your data"
subtitle: "SISBID 2016"
author: "Di Cook (dicook@monash.edu, @visnut); Heike Hofmann (heike.hofmann@gmail.com, @heike_hh)"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    transition: default
    widescreen: true
css:
  styles.css
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.height = 4,
  fig.width = 8,
  fig.align = "center",
  cache = FALSE
)
```

```{r echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(lubridate)
library(scales)
library(readr)
library(ggmap)
library(HLMdiag)
library(RColorBrewer)
library(gridExtra)
library(broom)
```

## Your turn




## Resources

- [Cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/08/ggplot2-cheatsheet.pdf)
- [ggplot2: Elegant Graphics for Data Analysis, Hadley Wickham](http://ggplot2.org/book/), [web site](http://ggplot2.org)
- [R Graphics Cookbook, Winston Chang](http://www.cookbook-r.com/Graphs/)
- [Q/A site: http://stackoverflow.com](http://stackoverflow.com)
- [Dynamic Documents with R and knitr, Yihui Xie, ](http://yihui.name/knitr/)
- [Naomi Robbins, Creating More Effective Graphs](http://www.nbr-graphs.com)
- [Antony Unwin, Graphical Data Analysis with R](https://www.crcpress.com/Graphical-Data-Analysis-with-R/Unwin/9781498715232)

## Share and share alike

This work is licensed under the Creative Commons Attribution-Noncommercial 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc/ 3.0/us/ or send a letter to Creative Commons, 171 Second Street, Suite 300, San Francisco, California, 94105, USA.
